A Fully Abstract Game Semantics for Countable Nondeterminism

### 2018-01-29 cones/paper/cones.tex
Fixed error in the last section.

The second observation is that Lemma \ref{lem:soundness-lemma} can just as easily be applied to PCF contexts and its game semantics model $\G_{inn}$.  
There really is nothing to this observation: since any PCF context may be interpreted in Idealized Algol, then Lemma \ref{lem:soundness-lemma} applies to it.
So we have narrowed \ref{lem:soundness-lemma} down to a result about $\G_{inn}$.

@@

It is then possible to deduce versions of Lemma \ref{lem:consistency} and Proposition \ref{prop:adequacy} from Lemma \ref{lem:soundness-lemma} as we did above, using the details of the inclusion functor from $\G_{inn}$ and the definition of composition in $\G_{TO}$.  
We can then deduce soundness in the usual way.

\begin{theorem}[Soundness for nondeterministic PCF]
  Let $M,N\from T$ be two terms of PCF with countable nondeterminism.
  If $\deno M\oes \deno N$, then $M\emam N$, where the denotation $\deno -$ is taken in the category $\G_{TO}$.  
\end{theorem}

An alternative line of attack is to first build a copy of our model of Idealized Algol with countable nondeterminism inside $\G_{TO}$, using our argument to prove soundness and adequacy, and then restrict down to nondeterministic PCF.

>>>>

We want a version of Lemma \ref{lem:soundness-lemma} that can be applied to PCF contexts and the game semantics model $\G_{inn}$ of PCF.  
Clearly, the lemma as currently stated does not hold in $\G_{inn}$, since the strategies $\sigma$ that we define are not innocent in general.

Instead, we need to look at the combinatorial structure of the strategies.  
We observe that if $\sigma_M\from\oc\bN\to\bC$ is a strategy, then there exists a deterministic strategy $\sigma\from\oc\bN$ such that $qa\in T_{\sigma;\sigma_M}$ if and only if there exists some $\s\in T_{\sigma_M}$ such that $\s\vert_{\bC}=qa$.  
In addition, there exists a deterministic strategy $\sigma\from\oc\bN$ such that $D_{\sigma;\sigma_M}\ne\emptyset$ if and only if $\sigma_M$ is not \emph{winning}; i.e., either $\sigma_M$ is not total or there exists some $\s\in T_{\sigma_M}$ with an infinite tail of moves in $\oc\bN$.  
We recover the following version of Lemma \ref{lem:soundness-lemma} which now applies to the interpretation of PCF in $\G_{inn}$:
\begin{lemma}
  Let $M=C[\wn]$ be a term of type $\com$.  
  Write $\sigma_M\from\oc\bN\to\bC$ for the denotation of the term $\lambda n.C[n]$.  
  \begin{itemize}
    \item $M\converges\skipp$ if and only if there exists $\s\in \sigma_M$ such that $\s\vert_{\bC}=qa$.  
    \item $M\mustconverge$ if and only if $\sigma_M$ is \emph{winning}.
  \end{itemize}
  \label{lem:pcf-soundness-lemma}
\end{lemma}

@@

It is then possible to deduce versions of Lemma \ref{lem:consistency} and Proposition \ref{prop:adequacy} from Lemma \ref{lem:pcf-soundness-lemma} as we did above for our model of Idealized Algol, using the details of the inclusion functor from $\G_{inn}$ and the definition of composition in $\G_{TO}$.  
We can then deduce soundness in the usual way.

It is worth noting an alternative line of attack, which is to first build a copy of our model of Idealized Algol with countable nondeterminism inside $\G_{TO}$, using our argument to prove soundness and adequacy, and then restrict down to nondeterministic PCF.
The approach above is more general, however, since it does not make use of the fact that Idealized Algol may be interpreted in $\G_{TO}$.
